Physcraper: a python package for continual update of evolutionary estimates using the Open Tree of Life

### 2020-07-24 2-body.Rmd
edit methods 3

  along with their corresponding identification numbers, i.e., their GenBank accession numbers, which will be used later to get the whole sequence from the GenBank database in a dedicated library within the "physcraper" folder, allowing for secondary analyses to run significantly faster.
- BLAST result sequences will be discarded if they fall outside a default minimum and maximum length of 80% and 120%, respectively, of the average length (gaps dropped) of sequences in the checked alignment.

@@

- New seqeunces that pass all filtering steps are added to the "csv" taxon summary file.

@@

- Physcraper uses the software MUSCLE [@edgar2004muscle] to perform DNA sequence alignments. Instructions on how to install all software dependencies used by Physcraper are provided in the documentation.
- The process to align new sequences consists of two steps. First, all new sequences are aligned using the default MUSCLE options.
- Second, a MUSCLE profile alignment is performed, in which the original alignment
  is used as a template to align the new sequences. This ensures that the final alignment
  follows the homology criteria established by the original alignment.
- The final alignment is not further processed by Physcraper. It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for alignment processing EXAMPLES OF SUCH TOOLS.
- While curating the alignment is a critical step, it is not a reproducible one. The main reason for its lack of reproducibility might be that it is hard to track changes made on the alignment. A form of version control, to register the differences between the alignment that was produced by the siftware and the manually curated aligment will be ideal. Versioning alignments and adding them to a  global database is a next step for us!
- Users may also use Physcraper to gather matched sequences only, and apply their own preferred alignment and phylogenetic inference methods.

>>>>

  along with their corresponding identification numbers, i.e., their GenBank accession numbers, which will be used later to get the whole sequence from the GenBank database into a dedicated library within the "physcraper" folder, allowing for secondary analyses to run significantly faster.
- BLAST result sequences will be discarded if they fall outside a default minimum and maximum length of 80% and 120%, respectively, of the average length without gaps of sequences in the checked alignment.

@@

- New sequences passing all filtering steps are added to the "csv" taxon summary file.

@@

- Physcraper uses the software MUSCLE [@edgar2004muscle] to perform DNA sequence alignments. Instructions on how to install all software dependencies used by Physcraper are provided in the documentation.
- The process to align new sequences consists of two steps. First, all new sequences are aligned using the default MUSCLE options.
- Second, a MUSCLE profile alignment is performed, in which the original alignment
  is used as a template to align the new sequences. This ensures that the final alignment
  follows the homology criteria established by the original alignment.
- The final alignment is not further processed by Physcraper. It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for automatic alignment processing [@castresana2000selection; @castresana2002gblocks].
- While curating the alignment is a critical step, it is not a reproducible one. The main reason for its lack of reproducibility might be that it is hard to track changes made on the alignment. A form of version control, to register the differences between the alignment that was produced by the software and the manually curated alignment will be ideal. Versioning alignments and adding them to a  global database is a next step for us!
- Users may also use Physcraper to only gather new GenBank sequences, to then apply their own preferred alignment and phylogenetic inference methods.

### 2020-07-24 docs/submission-mee-2020-07-24.tex
edit methods 3

\caption{The Physcraper software is fully described on its documentation website at https://physcraper.readthedocs.io/en/latest/, along with installation instructions, tutorials, examples and function usage documentation.}\label{fig:framework}
\end{figure}

@@

  along with their corresponding identification numbers, i.e., their GenBank accession numbers, which will be used later to get the whole sequence from the GenBank database in a dedicated library within the ``physcraper'' folder, allowing for secondary analyses to run significantly faster.

@@

  BLAST result sequences will be discarded if they fall outside a default minimum and maximum length of 80\% and 120\%, respectively, of the average length (gaps dropped) of sequences in the checked alignment.

@@

  New seqeunces that pass all filtering steps are added to the ``csv'' taxon summary file.

@@

\begin{itemize}
\tightlist
\item
  Physcraper uses the software MUSCLE (Edgar \protect\hyperlink{ref-edgar2004muscle}{2004}) to perform DNA sequence alignments. Instructions on how to install all software dependencies used by Physcraper are provided in the documentation.
\item
  The process to align new sequences consists of two steps. First, all new sequences are aligned using the default MUSCLE options.
\item
  Second, a MUSCLE profile alignment is performed, in which the original alignment
  is used as a template to align the new sequences. This ensures that the final alignment
  follows the homology criteria established by the original alignment.
\item
  The final alignment is not further processed by Physcraper. It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for alignment processing EXAMPLES OF SUCH TOOLS.
\item
  While curating the alignment is a critical step, it is not a reproducible one. The main reason for its lack of reproducibility might be that it is hard to track changes made on the alignment. A form of version control, to register the differences between the alignment that was produced by the siftware and the manually curated aligment will be ideal. Versioning alignments and adding them to a global database is a next step for us!
\item
  Users may also use Physcraper to gather matched sequences only, and apply their own preferred alignment and phylogenetic inference methods.
\end{itemize}

@@

>>>>

\caption{The Physcraper software is fully described on its documentation website at [physcraper.readthedocs.io](https://physcraper.readthedocs.io/en/latest/), along with installation instructions, tutorials, examples and function usage documentation.}\label{fig:framework}
\end{figure}

@@

  along with their corresponding identification numbers, i.e., their GenBank accession numbers, which will be used later to get the whole sequence from the GenBank database into a dedicated library within the ``physcraper'' folder, allowing for secondary analyses to run significantly faster.

@@

  BLAST result sequences will be discarded if they fall outside a default minimum and maximum length of 80\% and 120\%, respectively, of the average length without gaps of sequences in the checked alignment.

@@

  New sequences passing all filtering steps are added to the ``csv'' taxon summary file.

@@

\begin{itemize}
\tightlist
\item
  Physcraper uses the software MUSCLE (Edgar \protect\hyperlink{ref-edgar2004muscle}{2004}) to perform DNA sequence alignments. Instructions on how to install all software dependencies used by Physcraper are provided in the documentation.
\item
  The process to align new sequences consists of two steps. First, all new sequences are aligned using the default MUSCLE options.
\item
  Second, a MUSCLE profile alignment is performed, in which the original alignment
  is used as a template to align the new sequences. This ensures that the final alignment
  follows the homology criteria established by the original alignment.
\item
  The final alignment is not further processed by Physcraper. It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for automatic alignment processing (Castresana \protect\hyperlink{ref-castresana2000selection}{2000}, \protect\hyperlink{ref-castresana2002gblocks}{2002}).
\item
  While curating the alignment is a critical step, it is not a reproducible one. The main reason for its lack of reproducibility might be that it is hard to track changes made on the alignment. A form of version control, to register the differences between the alignment that was produced by the software and the manually curated alignment will be ideal. Versioning alignments and adding them to a global database is a next step for us!
\item
  Users may also use Physcraper to only gather new GenBank sequences, to then apply their own preferred alignment and phylogenetic inference methods.
\end{itemize}

@@

\leavevmode\hypertarget{ref-castresana2002gblocks}{}%
Castresana, J. (2002). GBLOCKS: Selection of conserved blocks from multiple alignments for their use in phylogenetic analysis. \emph{Version 0.91 b. Copyrighted by J. Castresana, EMBL}.

\leavevmode\hypertarget{ref-castresana2000selection}{}%
Castresana, J. (2000). Selection of conserved blocks from multiple alignments for their use in phylogenetic analysis. \emph{Molecular biology and evolution}, \textbf{17}, 540--552.
